---
title: "Weekly Meal Planner - User Guide and Benefits"
output:
  html_document:
    css: styles.css
---

### **🍽️ Intro**
The Weekly Meal Planner by A.M. is a powerful tool designed to help you effortlessly create a weekly meal schedule. Whether you're planning meals for yourself, your family, or an event, this tool streamlines the process and ensures that your meals are diverse and tailored to your preferences. Below, we'll guide you through how to use this tool and highlight its benefits.

#### **🛠️ Set Up the Environment**
   - Ensure you have R & RStudio installed on your computer
   - Save the R Notebook provided as "index.Rmd."
   - Install and load the necessary R packages: `tidyr`, `tidyverse`, and `googlesheets4`. You can do this using the `install.packages()` and `library()` functions in R.

#### **🌐 Authenticate with Google Sheets**
   - The tool utilizes Google Sheets as a data source. To access your Google Sheets, you'll need to authenticate your Google account with R. Use the `gs4_auth()` function, replacing `"YOUR_EMAIL@gmail.com"` with your Gmail address. This step grants the tool access to your Google Sheets.

#### **🍲 Prepare Your Meals in Google Sheets:**
   - Create a Google Sheet (or use an existing one) to list your favorite meals. You can access this sheet using this [link](https://docs.google.com/spreadsheets/d/13uEbDKVbkLz9PpYwXzOLZAgw_ZWppBm7pjnDbHnH9yA/edit#gid=0), and make a copy for yourself.
   - In the Google Sheet, include columns for "Meal_Name," "Type," "Prep_Time_min," "Recipe_Link," and "Effort." Fill in the details for each meal.

#### **💾 Export Google Sheet as CSV (Optional):**
   - You can export your Google Sheet as a CSV file if needed. However, this step is optional, as the tool can directly access Google Sheets.

#### **📊 Access Google Sheet Data:**
   - In the R Notebook, specify the URL of your Google Sheet using the `sheet_url` variable. Replace `"YOUR_GOOGLE_SHEET_URL"` with the actual URL of your Google Sheet.
   - Use the `googlesheets4::read_sheet()` function to load your Google Sheet data into R. The tool will automatically convert the data into a data frame for further processing.

#### **📅 Generate Your Meal Schedule:**
   - The tool generates a weekly meal schedule with three meals a day (breakfast, lunch, and dinner) for each day of the week (Monday to Sunday).
   - It ensures that meals are randomized and don't repeat within the same week.
   - The output provides details such as the meal name, preparation time, recipe link, and effort level.

### **🏆 Benefits**

**⏱️ Time-Saving.** This tool significantly reduces the time and effort required to create a weekly meal plan. Instead of manually organizing meals, the tool does it in seconds.

**🍽️ Diverse Meals.** By randomizing your meals, it helps ensure that you enjoy a variety of dishes throughout the week, preventing meal monotony.

**🔧 Customization.** You have full control over your meal choices. You can easily update your Google Sheet with new recipes or preferred meals, and the tool will adapt accordingly.

**🚀 Efficiency.** The tool is designed to be efficient and user-friendly. With just a few simple steps, you can generate a detailed meal plan tailored to your preferences.

**🌍 Access Anywhere.** Since your meal data is stored in a Google Sheet, you can access and edit it from anywhere, making meal planning a breeze even when you're on the go.

---

# Weekly Meal Planner

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(tidyr)
library(tidyverse)
library(googlesheets4)
library(knitr)
library(kableExtra)
gs4_auth(email = "mirza.mujanovic@gmail.com")
```

```{r,warning=FALSE,message=FALSE,echo=FALSE}
# Manual import
data <- read.csv("Meal Planner - Calendar Week 35.csv",header=TRUE)
#or via Google Sheet package
sheet_url <- "https://docs.google.com/spreadsheets/d/13uEbDKVbkLz9PpYwXzOLZAgw_ZWppBm7pjnDbHnH9yA/edit#gid=1859631947"
#data<- suppressMessages(googlesheets4::read_sheet(sheet_url,sheet = 1)) # IMPORTANT: sheet = 1 means it takes first sheet tab.
data$type <- tolower(data$type) #lower-case the type of the meal
```


```{r, warning=FALSE,message=FALSE,echo=FALSE}
# New data frame
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
sample_size <- length(days)

# Randomization
breakfast <- data %>% 
  filter(type=="breakfast") %>%
  sample_n(size=sample_size, replace = FALSE) %>%
  add_column(days)

lunch <- data %>% 
  filter(type=="lunch") %>%
  sample_n(size=sample_size, replace = FALSE) %>%
  add_column(days)

dinner <- data %>% 
  filter(type=="dinner") %>%
  sample_n(size=sample_size, replace = FALSE) %>%
  add_column(days)
```

```{r, warning=FALSE,message=FALSE,echo=FALSE}
# Final weekly meal schedule
schedule <- rbind.data.frame(breakfast, lunch, dinner)
result_table <- suppressMessages(schedule %>%
  group_by(days, type) %>%
  summarise(meal, prep_time_min, recipe_link, effort) %>%
  arrange(match(days, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))))

# Create a nicely formatted table
styled_table <- result_table %>%
  kable(format = "html",
        escape = FALSE,
        col.names = c("Day", "Type", "Meal", "Prep Time (min)", "Recipe Link", "Effort Level")) %>%
  kable_styling("striped", full_width = FALSE) %>%
  column_spec(1, bold = TRUE)  # Apply bold formatting to the first column (days)

# Create a nicely formatted table with adjusted column widths
styled_table <- result_table %>%
  kable(format = "html",
        escape = FALSE,
        align='l',
        col.names = c("Day", "Type", "Meal", "PrepTime (min)", "Recipe Link", "Effort Level")) %>%
  kable_styling("striped", full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%  # Apply bold formatting to the first column (days)
  column_spec(2, width = "70px") %>%  # Adjust the width of the "Type" column
  column_spec(3, width = "150px") %>%  # Adjust the width of the "Meal" column
  column_spec(4, width = "200px") %>%  # Adjust the width of the "Prep Time (min)" column
  column_spec(5, width = "200px") %>%  # Adjust the width of the "Recipe Link" column
  column_spec(6, width = "100px")     # Adjust the width of the "Effort Level" column

# Print the table with the specified styling
styled_table

```


```{r, echo=FALSE, warning=FALSE,message=FALSE,echo=FALSE}
# Saving to Google Sheet 
# Run only once!
#uppressMessages(sheet_write(data = result_table,
 #           ss = sheet_url,
  #          sheet = "Calendar Week 35")) # sheet = the week 
```

## R code
```{r, results='hide',eval=FALSE}
# Packages
library(tidyr)
library(tidyverse)
library(googlesheets4)
library(knitr)
library(kableExtra)
gs4_auth(email = "YOUR_GOOGLE_EMAIL") # INSERT YOUR G-MAIL ASSOCIATED WITH YOUR GOOGLE SHEET ACCOUNT

# Manual import
data <- read.csv("Meal Planner - Calendar Week 35.csv",header=TRUE)
#or via Google Sheet package
sheet_url <- "YOUR_GOOGLE_SHEET_URL"
#data<- suppressMessages(googlesheets4::read_sheet(sheet_url,sheet = 1)) # IMPORTANT: sheet = 1 means it takes first sheet tab.
data$type <- tolower(data$type) #lower-case the type of the meal

# New data frame with days
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
sample_size <- length(days)

# Randomization per meal type
breakfast <- data %>% 
  filter(type=="breakfast") %>%
  sample_n(size=sample_size, replace = FALSE) %>%
  add_column(days)

lunch <- data %>% 
  filter(type=="lunch") %>%
  sample_n(size=sample_size, replace = FALSE) %>%
  add_column(days)

dinner <- data %>% 
  filter(type=="dinner") %>%
  sample_n(size=sample_size, replace = FALSE) %>%
  add_column(days)

# Final weekly meal schedule
schedule <- rbind.data.frame(breakfast, lunch, dinner)
result_table <- suppressMessages(schedule %>%
  group_by(days, type) %>%
  summarise(meal, prep_time_min, recipe_link, effort) %>%
  arrange(match(days, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))))

# Create a nicely formatted table with adjusted column widths
styled_table <- result_table %>%
  kable(format = "html",
        escape = FALSE,
        align='l',
        col.names = c("Day", "Type", "Meal", "PrepTime (min)", "Recipe Link", "Effort Level")) %>%
  kable_styling("striped", full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%  # Apply bold formatting to the first column (days)
  column_spec(2, width = "70px") %>%  # Adjust the width of the "Type" column
  column_spec(3, width = "150px") %>%  # Adjust the width of the "Meal" column
  column_spec(4, width = "200px") %>%  # Adjust the width of the "Prep Time (min)" column
  column_spec(5, width = "200px") %>% # column # Adjust the width of the "Recipe Link" column
  column_spec(6, width = "100px")     # Adjust the width of the "Effort Level" 

# Print the table with the specified styling
styled_table

# Saving to the same Google Sheet you used above
SuppressMessages(sheet_write(data = result_table,
           ss = sheet_url,
           sheet = "Calendar Week 35")) # sheet = the week number

```

#### For more projects and information about me, check out [my website](https://mirzamujanovic.github.io/website/).